---
output:
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```


# Stress-genes features


*Objective*: With different sets of stress-responsive genes check their features and rank them by importance for predicting stress-responsive genes.

## Quality control: Fold change comparison with previous data

Execute the script that process raw data from published articles and put it in a folder ready for the comparing fold changes between experiments. Then prepare the data for plotting

```{r, eval=TRUE}

# Execute the script for processing stress responsive files from Gasch and tillings
source("scr/previousResponsive.R")

# Read the data

tilings = read.delim("data/derived_data/0.4M_NaCl_FC_15min_responsive_Tilings.txt", stringsAsFactors = F)
gasch = read.delim("data/derived_data/1M_Sorbitol_15min_FC_Gasch.txt", stringsAsFactors = F)
RNAseq = read.delim("results/S_cerevisiae/DESeq2/S_cerevisiae_merged_salt_lfc1_p05_results.tsv", stringsAsFactors = F)

# Tilings annotation has two ORFs sometimes, we keep the first.
tilings$name2 = unlist(lapply(strsplit(tilings$name, ","), function(x) return(x[1])))
# Remove duplicates
tilings = tilings[!duplicated(tilings$name2),]

# Common IDS between the three datasets
commonIDS = Reduce(intersect, list(tilings$name2, gasch$UID, RNAseq$ensembl_gene_id))

# Prepare for plotting
tilingsC = subset(tilings, name2 %in% commonIDS)
gaschC = subset(gasch, UID %in% commonIDS)
RNAseqC = subset(RNAseq, ensembl_gene_id %in% commonIDS)

tilingsC = tilingsC[order(tilingsC$name2),]
gaschC = gaschC[order(gaschC$UID),]
RNAseqC = RNAseqC[order(RNAseqC$ensembl_gene_id),]

fcPrevPlot = data.frame(ensembl_gene_id = RNAseqC$ensembl_gene_id,
           RNASeq = RNAseqC$log2FoldChange, 
           Gasch = gaschC$X1M.sorbitol...15.min,
           Tilings = tilingsC$wt.15..0.4M...wt.0, 
           stringsAsFactors = F)

write.table(x = fcPrevPlot, 
            file = "results/S_cerevisiae/fcComparison_previousData.tab", 
            sep = "\t", 
            quote = F, 
            row.names = F)
```


## Functional enrichment analysis

```{r}
dir.create("results/S_cerevisiae/EnrichmentMap", showWarnings = FALSE)
# Processing the data for functional enrichment analysis with gprofiler and enrichmentmap in cytoscape
up <- read.delim("results/S_cerevisiae/responsive_genes/responsive_genes_salt_lfc1_up_merged.tsv", stringsAsFactors = FALSE)
write.table(x = paste(up$ensembl_gene_id, collapse = " "), file = "results/S_cerevisiae/EnrichmentMap/gProfiler_format_upregulated.txt", row.names = FALSE, col.names = FALSE, quote = FALSE)

down <- read.delim("results/S_cerevisiae/responsive_genes/responsive_genes_salt_lfc1_down_merged.tsv", stringsAsFactors = FALSE)
write.table(x = paste(down$ensembl_gene_id, collapse = " "), file = "results/S_cerevisiae/EnrichmentMap/gProfiler_format_downregulated.txt", row.names = FALSE, col.names = FALSE, quote = FALSE)
```


## Annotations and FASTA sequences


* Download UTRs annotation (Nagalakshmi et al., 2008) from SGD yeast browser.

* Parse UTRs annotations to eliminate 5'UTR with the warning: Potential AUG annotation error. This UTRs conflict with CDS annotation and ended up located inside CDS.

* Eliminate YLL054C 3'UTR that is annotated inside CDS.

* Process annotations

* Concatenate annotations

* Sort them


* Obtain FASTA sequences with getFASTA using genome FASTA sequence from [ENSEMBL](ftp://ftp.ensembl.org/pub/release-89/fasta/saccharomyces_cerevisiae/dna/Saccharomyces_cerevisiae.R64-1-1.dna.toplevel.fa.gz)

* Parse in R to match the desired format and merge with [CDS sequences](ftp://ftp.ensembl.org/pub/release-89/fasta/saccharomyces_cerevisiae/cds/Saccharomyces_cerevisiae.R64-1-1.cds.all.fa.gz)


```{r, eval=TRUE}
# Execute the script that perform all the proposed steps
system2(command = "chmod", args = "a+x scr/main_annotation_processing.sh")
system2("./scr/main_annotation_processing.sh")
```


## Collect sequence data

In the form of 5'utr, cds, 3'utr

```{r, eval = TRUE}
library(biomartr)

# Read CDS downloaded from Ensembl ftp (cds release-92)
sc_CDS <- read_cds(file= "data/original_data/Saccharomyces_cerevisiae.R64-1-1_release-89.cds.all.fa.gz", 
                   obj.type = "Biostrings")

# Create a data frame from CDS sequences
genesCDS = data.frame(geneName = unlist(lapply(strsplit(names(sc_CDS), " "), function(x) return(x[1]))), 
                      CDS = sc_CDS, 
                      row.names = NULL, 
                      stringsAsFactors = F)

# Read UTRs from Nagalaskshmi et al., 2008
genesUTRs = read.delim("data/derived_data/Saccharomyces_cerevisiae.R64-1-1.89_w_UTRs_sequences.tab", 
                     stringsAsFactors = F, 
                     header = FALSE, 
                     col.names = c("geneFeature", "sequence"))

# Separate into 5'utr and 3'utr and in the desired format
genesUTRs$which_utr = unlist(lapply(strsplit(genesUTRs$geneFeature, ":"), function(x) return(x[1])))
genesUTRs$geneName = unlist(lapply(strsplit(genesUTRs$geneFeature, ":"), function(x) return(x[2])))
genes5UTR = subset(genesUTRs, which_utr == "5UTR", c(geneName, sequence))
colnames(genes5UTR) = c("geneName", "UTR5")
genes3UTR = subset(genesUTRs, which_utr == "3UTR", c(geneName, sequence))
colnames(genes3UTR) = c("geneName", "UTR3")

# Merge data frames
geneswUtrs = Reduce(function(x,y) merge(x,y, by = "geneName", all = TRUE), list(genes5UTR, genesCDS, genes3UTR))
# Remove gene that only has 3UTR due to annotation error
geneswUtrs = geneswUtrs[geneswUtrs$geneName != "YDL038C",]
# Genename column as row.names
row.names(geneswUtrs) = geneswUtrs$geneName
geneswUtrs = geneswUtrs[,-1]
```

## Length

```{r, eval=TRUE}
# lapply gives the result to a list
# with do.call, as.data.frame and cbind
# ordering as a data frame, specifying
# row.names from geneswUTRs
lengthGeneswUtrs <- as.data.frame(do.call(cbind, lapply(geneswUtrs, nchar)), 
                   row.names = row.names(geneswUtrs))
colnames(lengthGeneswUtrs) <- paste0("length_",colnames(lengthGeneswUtrs))
```

## GC content comparison

```{r, eval=TRUE}
library(stringr)
gcGeneswUtrs <- as.data.frame(do.call(cbind, lapply(geneswUtrs, function(x) (str_count(x, "G") + str_count(x, "C"))/ nchar(x))), 
                   row.names = row.names(geneswUtrs))
colnames(gcGeneswUtrs) <- paste0("gc_",colnames(gcGeneswUtrs))
```


## Codon Pair Bias

Computed based on: http://science.sciencemag.org/content/sci/suppl/2008/06/26/320.5884.1784.DC1/Coleman.SOM.pdf

```{r, eval=TRUE}
source("functions/CodonPairBias.R")
cpbGeneswUtrs <- CodonPairBias(genetable = geneswUtrs)
cpbGeneswUtrs <- data.frame(cpbGeneswUtrs, row.names = 1, stringsAsFactors = F)
```


## Structure: PARS score

PARS score object generated by René from [PARS publication data](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3847670/). Here we just consider the mean PARS score per gene.

```{r, eval = TRUE}
load("data/original_data/PARS_objects_Pablo.RData", verbose = T)
# PARS summary table use at it is
parsGeneswUtrs <- as.data.frame(parsSummaryTable[,c("PARS_UTR5", "PARS_CDS", "PARS_UTR3")])
```


## Prionic properties
Prion Like Amino Acid Composition as precomputed in the [PLAAC page](http://plaac.wi.mit.edu/Scer-all-proteins-2014-05-17.xls).

```{r, eval = TRUE}
plaac <- read.delim("data/original_data/plaac_yeast.txt", stringsAsFactors = F)
plaacGeneswUtrs <- data.frame(plaac[,c("SEQid", "LLR")], row.names = 1, stringsAsFactors = F)
```

## Gene expression variability computed from scRNA-seq
DM and residual variance computed from single cell data
```{r, eval = TRUE, warning=FALSE}

source("scr/yscGeneExsVar.R")

noise <- read.delim("results/S_cerevisiae/expsVariability_DM_resVar.tab", stringsAsFactors = F)
# First element of composed ORFs as name of the ORF
noise$name.first <- unlist(lapply(strsplit(noise$name, ","), function(x) return(x[1])))
# Deal with duplicated gene names (n = 8) that appear after removing first element of composed ORFs
dIds <- duplicated(noise$name.first)
noise$name.first.wd <- noise$name.first 
noise$name.first.wd[dIds] <- paste(noise$name.first.wd[dIds], "_2", sep = "")

noiseGeneswUtrs <- data.frame(noise[,c("name.first.wd", "DM", "vst_residual_variance")], row.names = 1, stringsAsFactors = F)
```

## Number of TF regulating a gene
Data downloaded from [YEASTRACT database]

```{r, eval = TRUE}
# Read data generated in http://kaph.inesc-id.pt/formgenerateregulationmatrix.php using Documented (DNA plus Expression), all ORFs, all TFs
regulatoryAssociations <- read.delim("data/original_data/RegulationTwoColumnTable_Documented_20201021_749_464933261.tsv", stringsAsFactors = F, sep = ";")

# Format data to have the gene and the transcription factors that regulate it
tfByGene <- with(regulatoryAssociations, split(TF, Target))

# Remove the only two genes that have a duplicated TF (MAL61 and MAL62)
tfByGene <- tfByGene[lapply(lapply(tfByGene, duplicated), sum) == 0] 
# Number of TF that regulates a gene
nTfByGene <- lapply(tfByGene, length)

# Generate a data frame
nTfByGene.df <- plyr::ldply(nTfByGene)
colnames(nTfByGene.df) <- c("gene", "nTf")
nTfGeneswUtrs <- data.frame(nTfByGene.df, row.names = 1, stringsAsFactors = F)

# nTFs is a mix of ensembl gene ids and gene external symbols
library(biomaRt)
sc <- useMart("ensembl", dataset = "scerevisiae_gene_ensembl", host = "http://may2017.archive.ensembl.org")
ensembl <- getBM(filters = "external_gene_name", 
                 attributes = c("ensembl_gene_id", "external_gene_name"),
                 values = row.names(nTfGeneswUtrs), 
                 mart = sc)

nTfGeneswUtrs$ensembl_gene_id <- ensembl[match(row.names(nTfGeneswUtrs), ensembl$external_gene_name),"ensembl_gene_id"]
nTfGeneswUtrs$ensembl_gene_id <- ifelse(is.na(nTfGeneswUtrs$ensembl_gene_id), yes = row.names(nTfGeneswUtrs), no = nTfGeneswUtrs$ensembl_gene_id)
nTfGeneswUtrs <- data.frame(nTfGeneswUtrs, row.names = 2, stringsAsFactors = F)
```

## Half-life

Time of an RNA to reduce its number to half. From [eLIFE article](https://elifesciences.org/articles/32536): Non-invasive measurement of mRNA decay reveals translation initiation as the major determinant of mRNA stability.

```{r, eval = TRUE}
# Read data downloaded from eLife article (https://elifesciences.org/articles/32536)
halfLife <- read.delim("data/original_data/elife-32536-fig1-data2-v4_halfLife_yeast_replicates.txt", stringsAsFactors = F)

# Data is two replicates
# Compute the mean, if only one value use this value
halfLife$halfLife <- rowMeans(halfLife[,3:4], na.rm = T)
# Substitue NaN by NA
halfLife[is.nan(halfLife$halfLife), "halfLife"] <- NA

# Now is normal, try to do the plot using both approaches and for the model maybe take the log2
halfLifeGeneswUtrs <- data.frame(halfLife[,c("gene_id", "halfLife")], stringsAsFactors = F, row.names = 1)
```

## Extra features from Koch et al., 2012 Genome biology

```{r, eval=TRUE}
# Read data, then merge
## Small warning: Only taking into account genes that are in both lists and could be merged
## In addition, some external gene identifiers of the table are ensembl gene ids
## Furthermore, there are official gene symbols that in the downloaded table are aliases or even OCT1 is transformed to 1-Oct due to excel
kochSupTable <- read.delim("data/original_data/Koch_2012_Genome_Biology_AF3.txt", 
                           stringsAsFactors = F)

# Change NaN to NA (from stack exchange: https://stackoverflow.com/questions/18142117/how-to-replace-nan-value-with-zero-in-a-huge-data-frame)
# Note that it is not required to call is.nan.data.frame due to method dispatch
is.nan.data.frame <- function(x)
do.call(cbind, lapply(x, is.nan))

kochSupTable[is.nan(kochSupTable)] <- NA
kochSupTable <- kochSupTable[!duplicated(kochSupTable$ensembl_gene_id),]

# Warning: The subseting is done by indices so if changed do it carefully
kochGeneswUtrs <- data.frame(kochSupTable[,2:19], stringsAsFactors = F, row.names = 1) 
# Remove length protein and copy number
kochGeneswUtrs <- kochGeneswUtrs %>%
  dplyr::select(-c(Copy_number, Protein_length))
```


## Gathering the data, exploratory and comparative analysis

Gathering features

```{r, eval= TRUE}
# Doing the gather of the data frames in an automatic way
properties.dataframes <- grep("GeneswUtrs", ls(), value = T)
properties.list <- lapply(properties.dataframes, get)
properties.list <- lapply(properties.list, function(x) data.frame(x, ensembl_gene_id = row.names(x)))
names(properties.list) <- properties.dataframes
# Creating an empty data frame with the identifieres of the 7126 genes annotation
# In this way we ensure to have a consistent annotation with the genes that match
properties.df <- data.frame("ensembl_gene_id" = row.names(geneswUtrs), stringsAsFactors = F)
print("Merging the following properties into a single data frame:")
for (i in names(properties.list)){
  property.name <- unlist(strsplit(i, "GeneswUtrs"))
  print(property.name)
  property <- properties.list[[i]]
  properties.df <- merge(properties.df, property, by = "ensembl_gene_id", all.x = T)
}
# Count number of complete observations for each property
nObservations <- plyr::ldply(colSums(!sapply(properties.df[,-1], is.na)))
colnames(nObservations) <- c("Property", "Observations")
write.table(x = nObservations, file = "results/S_cerevisiae/numberOfObservations_per_feature.tab", sep = "\t", quote = F, row.names = F)
```

Joining with the stress consensus classification

```{r, eval = TRUE}
# Stress responsive genes determined in Stress gene consensus project
stressGenes <- read.delim("results/S_cerevisiae/stressConsensusTable.tab", stringsAsFactors = F)
stress.properties.df <- merge(stressGenes, properties.df, by = "ensembl_gene_id", all.y = T)
# Order properties alphabetically
stress.properties.df <- stress.properties.df[,c(names(stress.properties.df)[1:4], sort(names(stress.properties.df[5:ncol(stress.properties.df)])))]
write.table(x = stress.properties.df, file = "results/S_cerevisiae/stressConsensus_Properties_Full_Table.tsv", sep = "\t", quote = F, row.names = F)
```

# Processing and selection of features. Modelling.
## Feature selection and data processing
```{r}
library(Hmisc)
library(nnet)

stress.properties.df <- read.delim("results/S_cerevisiae/stressConsensus_Properties_Full_Table.tsv", stringsAsFactors = F)
properties <- colnames(stress.properties.df)[5:ncol(stress.properties.df)]

prop.names.format <- read.delim("data/original_data/features_info.txt", stringsAsFactors = FALSE)

# Multinomial logistic regression requires low level of collinearity
# Here, we use pearson correlation to filter our variables that are linearly correlated
cor_all <- rcorr(as.matrix(stress.properties.df[, 5:ncol(stress.properties.df)]), type = "pearson")
# Write for plotting
save(cor_all, file = "data/rdata/pairwise_pearson_allFeatures.rda")

findCorrObs <- function(mat, cutoff = 0.35, numberIter){
 # Define final object for if
 finalFeatures <- c()
 for (i in 1:numberIter){
 # TRUE/FALSE matrix if value exceds correlation threshold
 fMat <- abs(mat$r) > cutoff
 # Number of observations matrix
 nMat <- mat$n
 finalFeatures <- unlist(lapply(row.names(fMat), function(feature){
    tmp = fMat[feature,]
    tmp = tmp[names(tmp) != feature]
    if(length(finalFeatures) > 0){
    tmp = tmp[names(tmp) %in% finalFeatures]
    }
    corrFeat = names(which(tmp))
    nObs = unlist(lapply(corrFeat, function(x){
      nMat[x,x]
    }))
    if(any(!nMat[feature, feature] >= nObs)){
    } else {
      feature
    }
  }))
 }
 return(finalFeatures)
}

keepFeatures <- findCorrObs(mat = cor_all, cutoff = 0.35, numberIter = 2) # with numberIter = 2 saturation is reached.

keepFeatures <- sub(x = keepFeatures, pattern = "Yeast_conservation", replacement = "Broad_conservation")
keepFeatures <- sub(x = keepFeatures, pattern = "vst_residual_variance", replacement = "DM")

keepFeatures <- keepFeatures[!keepFeatures %in% c("SM_fitness_defect", "PARS_UTR5", "PARS_UTR3", "Co.expression_degree")] # Filter them because they haveless than 4000 observations and affect too much final number of observations. Also filter co.expression degree (+ control)

# Order alphabetically
keepFeatures <- keepFeatures[order(keepFeatures)]

stress.properties.df.filtered <- stress.properties.df[,c(1:4,which(colnames(stress.properties.df) %in% keepFeatures))]
prop.names.format.filtered <- prop.names.format[prop.names.format$Property %in% keepFeatures,]
  
# Number of complete observations in total
table(rowSums(is.na(stress.properties.df.filtered[,5:ncol(stress.properties.df.filtered)])) == 0)

# Scale data
# Omit NA to have only complete observations
stress.properties.df.filtered.nar <- na.omit(stress.properties.df.filtered)
stress.properties.df.filtered.scaled <- cbind(stress.properties.df.filtered.nar[,1:4], apply(stress.properties.df.filtered.nar[,5:ncol(stress.properties.df.filtered.nar)], 2, scale))
```

## Multinomial logistic regression: Unresponsive, Downregulated, Upregulated

```{r}
# Multinomial logistic regression and save output for plotting
cond <- "salt"
depOverlap <- c("unresponsive", "downregulated", "upregulated")

# Raw data - Leave-one out (AIC)
for (j in c(0,400)){
  if(j == 0){
    raw.data <- stress.properties.df.filtered.nar
  } else {
    set.seed(12345)
    newIds <- c(sample(x = which(stress.properties.df.filtered.nar[,cond] == "unresponsive"), size = j), which(stress.properties.df.filtered.nar[,cond] != "unresponsive"))
    raw.data <- stress.properties.df.filtered.nar[newIds, ]
  }
    
  raw.data[,paste0(cond, "_2")] <- factor(raw.data[,cond])
  raw.data[,paste0(cond,"_2")] <- relevel(raw.data[,paste0(cond,"_2")], ref = "unresponsive")
  # Formula    
  raw.f <- as.formula(paste0(cond, "_2", " ~ ", paste0(keepFeatures, collapse = " + ")))
  # Multinomial test
  raw.test <- multinom(raw.f, data = raw.data)
  raw.test.res <- summary(raw.test)
  # Compute z-score and p-value and write output to a table
  z <- raw.test.res$coefficients/raw.test.res$standard.errors
  p <- (1 - pnorm(abs(z), 0, 1)) * 2
  raw.test.out <- as.data.frame(rbind(raw.test.res$coefficients, raw.test.res$standard.errors, p))
  row.names(raw.test.out)[grep(".1", row.names(raw.test.out))] <- gsub(".1", "_standardErrors", row.names(raw.test.out)[grep(".1", row.names(raw.test.out))], fixed = T)
  row.names(raw.test.out)[grep(".2", row.names(raw.test.out))] <- gsub(".2", "_pValue", row.names(raw.test.out)[grep(".2", row.names(raw.test.out))], fixed = T)
  write.table(x = raw.test.out,
                  file = paste0("results/S_cerevisiae/multinomialTestSummary_raw_",cond, "_random_", j, ".tab"), 
                  quote = F,
                  sep = "\t",
                  col.names = NA)
  
  # AIC
  raw.AIC.full <- raw.test$AIC
  raw.AIC.leave.one <- lapply(keepFeatures, function(x){
      raw.f.leave.one <- paste0(cond, "_2", " ~ ", paste0(grep(x, keepFeatures, invert = T, value = T), collapse = " + "))
      raw.test.leave.one <- multinom(raw.f.leave.one, data = raw.data)
      raw.test.leave.one$AIC
  })
  names(raw.AIC.leave.one) <- keepFeatures
  raw.AIC.leave.one$full <- raw.AIC.full
  raw.AIC.df <- plyr::ldply(raw.AIC.leave.one, .id = "Property")
  colnames(raw.AIC.df)[2] = "AIC"
  # Correct manually variables for the correct format
  raw.AIC.df$Property_Formatted <- paste0(" ", prop.names.format[match(raw.AIC.df$Property, prop.names.format$Property), "Property_formatted_short"], " ")
  raw.AIC.df[raw.AIC.df$Property == "full", "Property_Formatted"] <- " Full Model "
  # Save data frame from latter plotting
  write.table(raw.AIC.df, paste0("results/S_cerevisiae/dataAIC_", cond, "_random_", j,".tsv"), sep = "\t", quote = F, row.names = F)
  
  # Scaled data - Coefficient of regression comparison
  if(j == 0){
    scaled.data <- stress.properties.df.filtered.scaled
  } else {
    set.seed(12345)
    newIds <- c(sample(x = which(stress.properties.df.filtered.scaled[,cond] == "unresponsive"), size = j), which(stress.properties.df.filtered.scaled[,cond] != "unresponsive"))
    scaled.data <- stress.properties.df.filtered.scaled[newIds, ]
  }
    scaled.data[,paste0(cond, "_2")] <- factor(scaled.data[,cond])
    scaled.data[,paste0(cond,"_2")] <- relevel(scaled.data[,paste0(cond,"_2")], ref = "unresponsive")
    scaled.f <- as.formula(paste0(cond, "_2", " ~ ", paste0(keepFeatures, collapse = " + ")))
    # Multinomial test
    scaled.test <- multinom(scaled.f, data = scaled.data)
    scaled.test.res <- summary(scaled.test)
    # Compute z-score and p-value and write output to a table
    z <- scaled.test.res$coefficients/scaled.test.res$standard.errors
    p <- (1 - pnorm(abs(z), 0, 1)) * 2
    scaled.test.out <- as.data.frame(rbind(scaled.test.res$coefficients, scaled.test.res$standard.errors, p))
    row.names(scaled.test.out)[grep(".1", row.names(scaled.test.out))] <- gsub(".1", "_standardErrors", row.names(scaled.test.out)[grep(".1", row.names(scaled.test.out))], fixed = T)
    row.names(scaled.test.out)[grep(".2", row.names(scaled.test.out))] <- gsub(".2", "_pValue", row.names(scaled.test.out)[grep(".2", row.names(scaled.test.out))], fixed = T)
    write.table(x = scaled.test.out,
                file = paste0("results/S_cerevisiae/multinomialTestSummary_scaled_", cond,  "_random_", j, ".tab"), 
                quote = F,
                sep = "\t",
                col.names = NA)
  
  # Accuracy of multinomial logistic regression: Split the data into training and validation set
  # Training Set : Validation Set = 70 : 30 (random)
  set.seed(12345)
  train <- sample(nrow(raw.data), 0.7*nrow(raw.data))
  TrainSet <- raw.data[train,]
  ValidSet <- raw.data[-train,]
  # Create a Multinomial Logistic Regression model with default parameters
  model1 <- multinom(raw.f, data = TrainSet)
  # Predicting on Validation set
  predValid <- predict(model1, ValidSet, type = "class")
  # Checking classification accuracy
  mean(predValid == ValidSet[, paste0(cond, "_2")])
  # ROC curve assesment
  # Calculate the probability of new observations belonging to each class
  # prediction_for_roc_curve will be a matrix with dimensions data_set_size x number_of_classes
  prediction_for_roc_curve <- predict(model1, ValidSet, type ="prob")
  # Save for plotting
  save(prediction_for_roc_curve, ValidSet, file = paste0("data/rdata/roc_curve_multinomial_scerevisiae_random_",j,".rda"))
}
```

## Multinomial logistic regression: Unresponsive, Downregulated, Upregulated (+ coexpression degree)

```{r}
keepFeatures.c <- c(keepFeatures, "Co.expression_degree") 
# Order alphabetically
keepFeatures.c <- keepFeatures.c[order(keepFeatures.c)]

stress.properties.df.filtered.c <- stress.properties.df[,c(1:4,which(colnames(stress.properties.df) %in% keepFeatures.c))]
prop.names.format.filtered.c <- prop.names.format[prop.names.format$Property %in% keepFeatures.c,]
  
# Number of complete observations in total
table(rowSums(is.na(stress.properties.df.filtered.c[,5:ncol(stress.properties.df.filtered.c)])) == 0)

# Scale data
# Omit NA to have only complete observations
stress.properties.df.filtered.nar.c <- na.omit(stress.properties.df.filtered.c)

# Multinomial logistic regression and save output for plotting
cond <- "salt"
depOverlap <- c("unresponsive", "downregulated", "upregulated")

# Raw data - Leave-one out (AIC)
for (j in c(0,400)){
  if(j == 0){
    raw.data <- stress.properties.df.filtered.nar.c
  } else {
    set.seed(12345)
    newIds <- c(sample(x = which(stress.properties.df.filtered.nar.c[,cond] == "unresponsive"), size = j), which(stress.properties.df.filtered.nar.c[,cond] != "unresponsive"))
    raw.data <- stress.properties.df.filtered.nar.c[newIds, ]
  }
    
  raw.data[,paste0(cond, "_2")] <- factor(raw.data[,cond])
  raw.data[,paste0(cond,"_2")] <- relevel(raw.data[,paste0(cond,"_2")], ref = "unresponsive")
  # Formula    
  raw.f <- as.formula(paste0(cond, "_2", " ~ ", paste0(keepFeatures.c, collapse = " + ")))
  # Multinomial test
  raw.test <- multinom(raw.f, data = raw.data)
  raw.test.res <- summary(raw.test)
  # Compute z-score and p-value and write output to a table
  z <- raw.test.res$coefficients/raw.test.res$standard.errors
  p <- (1 - pnorm(abs(z), 0, 1)) * 2
  raw.test.out <- as.data.frame(rbind(raw.test.res$coefficients, raw.test.res$standard.errors, p))
  row.names(raw.test.out)[grep(".1", row.names(raw.test.out))] <- gsub(".1", "_standardErrors", row.names(raw.test.out)[grep(".1", row.names(raw.test.out))], fixed = T)
  row.names(raw.test.out)[grep(".2", row.names(raw.test.out))] <- gsub(".2", "_pValue", row.names(raw.test.out)[grep(".2", row.names(raw.test.out))], fixed = T)
  write.table(x = raw.test.out,
                  file = paste0("results/S_cerevisiae/multinomialTestSummary_raw_",cond, "_random_", j, "_coexpsDegree.tab"), 
                  quote = F,
                  sep = "\t",
                  col.names = NA)
  
  # AIC
  raw.AIC.full <- raw.test$AIC
  raw.AIC.leave.one <- lapply(keepFeatures.c, function(x){
      raw.f.leave.one <- paste0(cond, "_2", " ~ ", paste0(grep(x, keepFeatures.c, invert = T, value = T), collapse = " + "))
      raw.test.leave.one <- multinom(raw.f.leave.one, data = raw.data)
      raw.test.leave.one$AIC
  })
  names(raw.AIC.leave.one) <- keepFeatures.c
  raw.AIC.leave.one$full <- raw.AIC.full
  raw.AIC.df <- plyr::ldply(raw.AIC.leave.one, .id = "Property")
  colnames(raw.AIC.df)[2] = "AIC"
  # Correct manually variables for the correct format
  raw.AIC.df$Property_Formatted <- paste0(" ", prop.names.format[match(raw.AIC.df$Property, prop.names.format$Property), "Property_formatted_short"], " ")
  raw.AIC.df[raw.AIC.df$Property == "full", "Property_Formatted"] <- " Full Model "
  # Save data frame from latter plotting
  write.table(raw.AIC.df, paste0("results/S_cerevisiae/dataAIC_", cond, "_random_", j,"_coexpsDegree.tsv"), sep = "\t", quote = F, row.names = F)
}
```

## Logistic regression: Hog1 dependent vs Hog1 independent genes

```{r, eval = TRUE}
# Logistic regression in Hog1 dependency genes
cond <- "hog1DepSaltExt"
depOverlap <- c("unresponsive","induced_Hog1_independent", "induced_Hog1_dependent", "repressed_Hog1_independent" ,"repressed_Hog1_dependent")
    
# Logistic regression
# Raw data - Leave-one out - AIC
raw.data <- stress.properties.df.filtered.nar
raw.data[,paste0(cond, "_2")] <- factor(raw.data[,cond])

# Scaled data - Model coefficients
scaled.data <- stress.properties.df.filtered.scaled
scaled.data[,paste0(cond, "_2")] <- factor(scaled.data[,cond])
  

for (direction in c("upregulated", "downregulated")){
  print(direction)
  print("Raw")
  raw.data.s <- raw.data[raw.data[,paste0(cond, "_2")] %in% c(paste0(direction, "_Hog1_independent"), paste0(direction, "_Hog1_dependent")),]
  raw.data.s[, paste0(cond, "_2")] <- droplevels(raw.data.s$hog1DepSaltExt_2)
  raw.data.s[, paste0(cond, "_2")] <- relevel(raw.data.s[,paste0(cond,"_2")], ref = paste0(direction, "_Hog1_independent"))
  raw.f <- as.formula(paste0(cond, "_2", " ~ ", paste0(keepFeatures, collapse = " + ")))
  raw.test.s <- glm(raw.f, data = raw.data.s, family = binomial(link = "logit"))
  raw.test.res.s <- summary(raw.test.s)
  write.table(raw.test.res.s$coefficients, 
              paste0("results/S_cerevisiae/logisticTestSummary_raw_", direction,"_", cond, ".tab"),
              quote = F,
              sep = "\t",
              col.names = NA)
  
  # AIC leave-one-out
  raw.AIC.full <- raw.test.s$aic
  raw.AIC.leave.one <- lapply(keepFeatures, function(x){
    raw.f.leave.one <- as.formula(paste0(cond, "_2", " ~ ", paste0(grep(x, keepFeatures, invert = T, value = T), collapse = " + ")))
    raw.test.leave.one <- glm(raw.f.leave.one, data = raw.data.s, family =  binomial(link = "logit"))
    raw.test.leave.one$aic
  })
  names(raw.AIC.leave.one) <- keepFeatures
  raw.AIC.leave.one$full <- raw.AIC.full
  raw.AIC.df <- plyr::ldply(raw.AIC.leave.one, .id = "Property")
  colnames(raw.AIC.df)[2] = "AIC"
  # Correct manually variables for the correct format
  raw.AIC.df$Property_Formatted <- paste0(" ", prop.names.format[match(raw.AIC.df$Property, prop.names.format$Property), "Property_formatted_short"], " ")
  raw.AIC.df[raw.AIC.df$Property == "full", "Property_Formatted"] <- " Full Model "
  # Save data frame from latter plotting
  write.table(raw.AIC.df, paste0("results/S_cerevisiae/dataAIC_", direction, "_", cond, ".tsv"), sep = "\t", quote = F, row.names = F)

  print("Scaled")
  # Scaled data - Coefficient of regression comparison
  scaled.data.s <- scaled.data[scaled.data[,paste0(cond, "_2")] %in% c(paste0(direction, "_Hog1_independent"), paste0(direction, "_Hog1_dependent")),]
  scaled.data.s[, paste0(cond, "_2")] <- droplevels(scaled.data.s$hog1DepSaltExt_2)
  scaled.data.s[, paste0(cond, "_2")] <- relevel(scaled.data.s[,paste0(cond,"_2")], ref = paste0(direction, "_Hog1_independent"))
  scaled.f <- as.formula(paste0(cond, "_2", " ~ ", paste0(keepFeatures, collapse = " + ")))
  # Multinomial test
  scaled.test <- glm(scaled.f, data = scaled.data.s, family = binomial(link = "logit"))
  scaled.test.res <- summary(scaled.test)
  write.table(x = scaled.test.res$coefficients,
              file = paste0("results/S_cerevisiae/logisticTestSummary_scaled_",direction, "_", cond, ".tab"), 
              quote = F,
              sep = "\t",
              col.names = NA)
  
  
  # Accuracy of logistic regression: Split the data into training and validation set
  # Training Set : Validation Set = 70 : 30 (random)
  set.seed(12345)
  train <- sample(nrow(raw.data.s), 0.7*nrow(raw.data.s), replace = F)
  TrainSet <- raw.data.s[train,]
  ValidSet <- raw.data.s[-train,]
  # Create a Logistic Regression model with default parameters (Careful:Doing it with multinom, since predict does not accept glm. Results are the same)
  model1 <- multinom(raw.f, data = TrainSet)
  # Predicting on Validation set
  predValid <- predict(model1, ValidSet, type = "class")
  # Checking classification accuracy
  mean(predValid == ValidSet[, paste0(cond, "_2")])                    
  # ROC curve assesment
  # Calculate the probability of new observations belonging to each class
  # prediction_for_roc_curve will be a matrix with dimensions data_set_size x number_of_classes
  prediction_for_roc_curve <- predict(model1, ValidSet, type ="prob")
  save(prediction_for_roc_curve, ValidSet, file = paste0("data/rdata/roc_curve_logistic_scerevisiae_", cond, "_", direction,".rda"))
}
```




## Random forest: Unresponsive, Downregulated, Upregulated

```{r, eval = TRUE}
library(randomForest)

cond <- "salt"
depOverlap <- c("unresponsive", "donwregulated", "upregulated")

for (j in c(0,400)){
  # Sample to have a more balanced set
  if (j == 0){
    raw.data <- stress.properties.df.filtered.nar
  } else if (j == 400){
    set.seed(12345)
    newIds <- c(sample(x = which(stress.properties.df.filtered.nar[,cond] == "unresponsive"), size = j), which(stress.properties.df.filtered.nar[,cond] != "unresponsive"))
    raw.data <- stress.properties.df.filtered.nar[newIds, ]
    } 
  raw.data[,paste0(cond, "_2")] <- factor(raw.data[,cond])
  raw.data[,paste0(cond,"_2")] <- relevel(raw.data[,paste0(cond,"_2")], ref = "unresponsive")
  raw.f <- as.formula(paste0(cond, "_2", " ~ ", paste0(keepFeatures, collapse = " + ")))
  
  # Whole dataset: Mean decrease in accuracy Random Forest
  set.seed(12345)
  fullModel1 <- randomForest(raw.f, data = raw.data, importance = T)
  write.table(importance(fullModel1), file = paste0("results/S_cerevisiae/modelImportance_randomForest", cond, "_random_",j,".tab"),
              quote = F, 
              sep = "\t", 
              row.names = T, 
              col.names = NA)
  
  
      
   # Training Set : Validation Set = 70 : 30 (random)
   set.seed(12345)
   train <- sample(nrow(raw.data), 0.7*nrow(raw.data), replace = F)
   TrainSet <- raw.data[train,]
   ValidSet <- raw.data[-train,]
   # Create a Random Forest model with default parameters
   set.seed(12345)
   model1 <- randomForest(raw.f, data = TrainSet, importance = T)
   # Predicting on Validation set
   predValid <- predict(model1, ValidSet, type = "class")
   # Checking classification accuracy
   mean(predValid == ValidSet[, paste0(cond, "_2")])
   # ROC curve assesment
   # Calculate the probability of new observations belonging to each class
   # prediction_for_roc_curve will be a matrix with dimensions data_set_size x number_of_classes
   prediction_for_roc_curve <- predict(model1, ValidSet, type ="prob")
   save(prediction_for_roc_curve, ValidSet, file = paste0("data/rdata/roc_curve_randomForest_scerevisiae_random_", j, ".rda"))
      
}
```

## Random forest: Hog1 dependent vs Hog1 independent genes

```{r, eval = TRUE}
# Random forest in Hog1 dependency genes
# Logistic regression in Hog1 dependency genes
cond <- "hog1DepSaltExt"
depOverlap <- c("unresponsive","induced_Hog1_independent", "induced_Hog1_dependent", "repressed_Hog1_independent" ,"repressed_Hog1_dependent")
    
# Random forest
# Raw data - Leave-one out - AIC
raw.data <- stress.properties.df.filtered.nar
raw.data[,paste0(cond, "_2")] <- factor(raw.data[,cond])


for (direction in c("upregulated", "downregulated")){
  print(direction)
  raw.data.s <- raw.data[raw.data[,paste0(cond, "_2")] %in% c(paste0(direction, "_Hog1_independent"), paste0(direction, "_Hog1_dependent")),]
  raw.data.s[, paste0(cond, "_2")] <- droplevels(raw.data.s$hog1DepSaltExt_2)
  raw.data.s[, paste0(cond, "_2")] <- relevel(raw.data.s[,paste0(cond,"_2")], ref = paste0(direction, "_Hog1_independent"))
  raw.f <- as.formula(paste0(cond, "_2", " ~ ", paste0(keepFeatures, collapse = " + ")))

  
  # Whole data: Mean decrease in accuracy
  set.seed(12345)
  fullModel1 <- randomForest(raw.f, data = raw.data.s, importance = T)
  write.table(importance(fullModel1), 
              file = paste0("results/S_cerevisiae/modelImportance_randomForest_",cond, "_", direction, ".tab"),
              quote = F, 
              sep = "\t", 
              row.names = T, 
              col.names = NA)  
 
  
  # Training Set : Validation Set = 70 : 30 (random)
  set.seed(12345)
  train <- sample(nrow(raw.data.s), 0.7*nrow(raw.data.s), replace = F)
  TrainSet <- raw.data.s[train,]
  ValidSet <- raw.data.s[-train,]
  # Create a Random Forest model with default parameters
  set.seed(12345)
  model1 <- randomForest(raw.f, data = TrainSet, importance = T)
  # Predicting on Validation set
  predValid <- predict(model1, ValidSet, type = "class")
  # Checking classification accuracy
  mean(predValid == ValidSet[, paste0(cond, "_2")])                    
  # ROC curve assesment
  # Calculate the probability of new observations belonging to each class
  # prediction_for_roc_curve will be a matrix with dimensions data_set_size x number_of_classes
  prediction_for_roc_curve <- predict(model1, ValidSet, type ="prob")
  save(prediction_for_roc_curve, ValidSet, file = paste0("data/rdata/roc_curve_randomForest_scerevisiae_", cond, "_", direction, ".rda"))
}

```

## Multinomial logistic regression: Univariate modelling (unresponsive, downregulated, upregulated) 

```{r}
cond <- "salt"
depOverlap <- c("unresponsive", "downregulated", "upregulated")
    
# Multinomial logistic regression: Univariate approach
# Raw data
for (j in c(0,400)){
  if(j == 0){
    raw.data <- stress.properties.df.filtered.nar
  } else {
    set.seed(12345)
    newIds <- c(sample(x = which(stress.properties.df.filtered.nar[,cond] == "unresponsive"), size = j), which(stress.properties.df.filtered.nar[,cond] != "unresponsive"))
    raw.data <- stress.properties.df.filtered.nar[newIds, ]
  }
    
raw.data[,paste0(cond, "_2")] <- factor(raw.data[,cond])
raw.data[,paste0(cond,"_2")] <- relevel(raw.data[,paste0(cond,"_2")], ref = "unresponsive")

# For each property, do a model and store the output (AIC, pseudoR2)
univariate.aic = lapply(keepFeatures, function(ppty){
      print(ppty)
      f <- as.formula(paste0(cond, "_2", " ~ ", ppty))
      raw.test <- multinom(f, data = raw.data)
      raw.aic = raw.test$AIC
      return(raw.aic)
})
    
names(univariate.aic) <- keepFeatures
# AIC  
univariate.aic = plyr::ldply(univariate.aic)
colnames(univariate.aic) = c("Property", "AIC")
univariate.aic$Property_Formatted <- paste0(" ", prop.names.format[match(univariate.aic$Property, prop.names.format$Property), "Property_formatted_short"], " ")
# Save data frame from latter plotting
write.table(univariate.aic, paste0("results/S_cerevisiae/dataAIC_univariate_", cond,"_random_",j, ".tsv"), sep = "\t", quote = F, row.names = F)
}

```

## Logistic regression: Univariate modelling (Hog1 dependent vs independent genes) 

```{r}
# Logistic regression in Hog1 dependency genes
cond <- "hog1DepSaltExt"
depOverlap <- c("unresponsive","induced_Hog1_independent", "induced_Hog1_dependent", "repressed_Hog1_independent" ,"repressed_Hog1_dependent")
    
# Logistic regression
# Raw data
raw.data <- stress.properties.df.filtered.nar
raw.data[,paste0(cond, "_2")] <- factor(raw.data[,cond])

for (direction in c("upregulated", "downregulated")){
  print(direction)
  raw.data.s <- raw.data[raw.data[,paste0(cond, "_2")] %in% c(paste0(direction, "_Hog1_independent"), paste0(direction, "_Hog1_dependent")),]
  raw.data.s[, paste0(cond, "_2")] <- droplevels(raw.data.s$hog1DepSaltExt_2)
  raw.data.s[, paste0(cond, "_2")] <- relevel(raw.data.s[,paste0(cond,"_2")], ref = paste0(direction, "_Hog1_independent"))
  
  # For each property, logistic regression and store AIC and pseudoR2
  univariate.aic = lapply(keepFeatures, function(ppty){
      print(ppty)
      f <- as.formula(paste0(cond, "_2", " ~ ", ppty))
      raw.test <- glm(f, data = raw.data.s, family = binomial(link = "logit"))
      raw.aic = raw.test$aic
      return(aic = raw.aic)
    })
  
  names(univariate.aic) <- keepFeatures
  # AIC  
  univariate.aic = plyr::ldply(univariate.aic)
  colnames(univariate.aic) = c("Property", "AIC")
  univariate.aic$Property_Formatted <- paste0(" ", prop.names.format[match(univariate.aic$Property, prop.names.format$Property), "Property_formatted_short"], " ")
  # Save data frame from latter plotting
  write.table(univariate.aic, 
              paste0("results/S_cerevisiae/dataAIC_univariate_", direction, "_", cond, ".tsv"), 
              sep = "\t", 
              quote = F, 
              row.names = F)
}
```



## TF motif enrichment: Upregulated Hog1 dependent vs independent genes

### Generate responsive genes lists for HOMER motif analysis

```{r, eval = TRUE}
stressGroupsHog1 = split(stress.properties.df$ensembl_gene_id, stress.properties.df$hog1DepSaltExt)
lapply(names(stressGroupsHog1)[3:5], function(grp){
  write.table(x = stressGroupsHog1[[grp]], file = paste0("data/derived_data/", grp, "_toGrep.txt"), row.names = F, quote = F, col.names = F)
})
```

### Execute HOMER analysis

```{r}
system(command = "scr/HOMER_analysis.sh")
```

