```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

# Human stress genes features

*Objective*: Investigate yeast stress features in humans.

```{r}
# Make folder for results
dir.create("results/human", showWarnings = F)
```

## Gene gathering

Experimentally defined osmostress-responsive genes with RNA-seq data in HeLa cells (100 mM NaCl, 3 hours)

```{r}
source("scr/human/genesGather.R")
```


## Properties gathering

First gather human IDs (one ENSG one ENST relationship). Using APPRIS

```{r}
source("scr/human/humanIdsGather.R")
```

Second, gather human properties and create a table with the correct format

```{r}
source("scr/human/propertiesGather.R")
```

## Feature modelling


```{r}
source("scr/human/modellingProperties.R")
```


## TCGA analysis of stress genes

```{r}
dir.create("data/original_data/tcga.rna", showWarnings = FALSE)
```

```{bash}
for i in ~/projects/public_data/2016_01_28_TCGA_RNAseq/*/*data.txt; do
       cancerType=`basename "$i" | awk '{split($0,a,"."); print a[1]}'`;
       mkdir -p data/original_data/tcga.rna/"$cancerType"
	     ln -sf "$i" data/original_data/tcga.rna/"$cancerType"/
	     ln -sf ~/projects/public_data/TCGA_2016_01_28/"$cancerType"/tumour/patient_annotation.txt \
		       data/original_data/tcga.rna/"$cancerType"/ 
done

```

```{r}
source("scr/human/tcgaRNAseq.R")
```

```{r}
source("scr/human/tcgaCN.R")
```


